Towards Proof-Reconstruction of Problems in Classical Propositional Logic in Agda

### 2017-09-05 paper/sections/introduction.tex
minor changes.

A proof reconstruction tool provides such an integration in one direction,
is a translator of the reply delivered by the prover into the formalism of
the proof assistant. Because the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the target, the
reconstruction turns out in a ``reverse engineering'' task.
Then, reconstructing a proof involves an in-depth understanding of the
algorithms in the \ATP and the specification logic in the \ITP.
What we need from the \ATP tools is a proof-object in a consistent format
to work with, that is, a full script describing step-by-step with
exhaustive details and without ambiguities, the table of the derivation
to get the actual proof. For problems in classical propositional logic
(henceforth \CPL), from a list of at least forty\footnote{\ATPs available
from the web service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just
a few provers were able to deliver proofs (e.g., \name{CVC4}~
\cite{Barrett2011}, \name{SPASS}, and \name{Waldmeister}~\cite{hillenbrand1997})
and a little bit less replied with a proof in a file format like
\TSTP~\cite{sutcliffe2004tstp} (e.g., \name{E}, \Metis,
\name{Vampire}, and \name{Z3}), \len{LFSC}~\cite{Stump2008}, or the
\len{SMT-LIB}~\cite{Bohme2011} format.

>>>>

A proof reconstruction tool provides such an integration translating 
the reply delivered by the prover into the formalism of
the proof assistant. Because the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the target, the
reconstruction turns out in a ``reverse engineering'' task.
Then, reconstructing a proof involves an in-depth understanding of the
algorithms in the \ATP and the specification logic in the \ITP.
What we need from the \ATP tools is a proof-object in a consistent format
to work with, that is, a full script describing step-by-step with
exhaustive details and without ambiguities, the table of the derivation
to get the actual proof. For problems in classical propositional logic
(henceforth \CPL), from a list of at least forty\footnote{\ATPs available
from the web service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just
a few provers were able to deliver proofs (e.g., \name{CVC4}~
\cite{Barrett2011}, \name{SPASS}, and \name{Waldmeister}~\cite{hillenbrand1997})
and a little bit less replied with a proof in a file format like
\TSTP~\cite{sutcliffe2004tstp} (e.g., \name{E}, \Metis,
\name{Vampire}, and \name{Z3}), \len{LFSC}~\cite{Stump2008}, or the
\len{SMT-LIB}~\cite{Bohme2011} format.

### 2017-09-05 paper/sections/related-work.tex
minor changes.

\name{Sledgehammer} is a tool for \name{Isabelle} proof assistant
~\cite{paulson1994isabelle} that provides a full integration of automatic
theorem provers including ATPs
(see, for example, \cite{meng2006automation}, \cite{blanchette2013extending}
and \cite{Fleury2014}) and\SMTsolvers (see, for instance,
\cite{hurlin07practical}, \cite{bohme2010},
\cite{blanchette2013extending}, and \cite{Fleury2014}) with
\name{Isabelle/HOL} \cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic. Sultana, Benzm{\"{u}}ller, and
Paulson~\cite{Een2004} integrates \name{Leo-II} and \name{Satallax}, two
theorem provers for high-order logic with \name{Isabelle/HOL} proposing a
modular proof reconstruction work flow.

\name{SMTCoq}~\cite{armand2011,Ekici2017} is a tool for the \name{Coq}
proof assistant \cite{coqteam} which provides a certified checker for
proof witness coming from the\SMTsolver \name{veriT} \cite{bouton2009}
and adds a new tactic named verit, that calls \name{veriT} on any
\name{Coq} goal. In \cite{bezem2002automated}, given a fixed but arbitrary
first-order signature, the authors transform a proof produced by the
first-order automatic theorem prover \name{Bliksem} \cite{deNivelle2003}
in a \name{Coq} proof term.

Hurd~\cite{Hurd1999} integrates the first-order resolution prover
\name{Gandalf} with \name{HOL}~\cite{norrish2007hol}, a high-order theorem
prover, following an LCF model implementing the tactic \verb!GANDALF_TAC!.
For \name{HOL Light}, a version of \name{HOL} but with a simpler logic core,
the\SMTsolver \name{CVC4} was integrated, and Kaliszyk and
Urban~\cite{kaliszyk2013} reconstruct proofs from different ATPs with their
\name{PRocH} tool, replaying the detailed inference steps from the ATPs with
internal inference methods implemented in \name{HOL Light}.

@@

Kanso and Setzer~\cite{kanso2016light} integrate \name{Z3} in \Agda
but they also integrate the propositional fragment of the \name{E} prover in
\cite{Kanso2012}. They cataloged these two integration as
\emph{Oracle and Reflection} and \emph{Oracle and Justification}, respectively.
Their integration with the \name{E} prover is the most related work with our
apporach found at the moment while we write this document. Besides the use
of the same ITP, crucial differences easily arise like decidability of
propositional logic using semantics that we do not take into account for the
reason we describe later on in section~\ref{sec:conclusions}.

>>>>

\name{Sledgehammer} is a tool for \name{Isabelle} proof assistant
~\cite{paulson1994isabelle} that provides a full integration of automatic
theorem provers including \ATPs
(see, for example, \cite{meng2006automation}, \cite{blanchette2013extending}
and \cite{Fleury2014}) and \SMT solvers (see, for instance,
\cite{hurlin07practical}, \cite{bohme2010},
\cite{blanchette2013extending}, and \cite{Fleury2014}) with
\name{Isabelle/HOL} \cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic. Sultana, Benzm{\"{u}}ller, and
Paulson~\cite{Een2004} integrates \name{Leo-II} and \name{Satallax}, two
theorem provers for high-order logic with \name{Isabelle/HOL} proposing a
modular proof reconstruction work flow.

\name{SMTCoq}~\cite{armand2011,Ekici2017} is a tool for the \name{Coq}
proof assistant \cite{coqteam} which provides a certified checker for
proof witness coming from the \SMT solver \name{veriT} \cite{bouton2009}
and adds a new tactic named verit, that calls \name{veriT} on any
\name{Coq} goal. In \cite{bezem2002automated}, given a fixed but arbitrary
first-order signature, the authors transform a proof produced by the
first-order automatic theorem prover \name{Bliksem} \cite{deNivelle2003}
in a \name{Coq} proof term.

Hurd~\cite{Hurd1999} integrates the first-order resolution prover
\name{Gandalf} with \name{HOL}~\cite{norrish2007hol}, a high-order theorem
prover, following an \abbre{LCF} model implementing the tactic \verb!GANDALF_TAC!.
For \name{HOL Light}, a version of \name{HOL} but with a simpler logic core,
the \SMT solver \name{CVC4} was integrated, and Kaliszyk and
Urban~\cite{kaliszyk2013} reconstruct proofs from different ATPs with their
\name{PRocH} tool, replaying the detailed inference steps from the ATPs with
internal inference methods implemented in \name{HOL Light}.

@@

Kanso and Setzer~\cite{kanso2016light} integrate \name{Z3} in \Agda
but they also integrate the propositional fragment of the \name{E} prover in
\cite{Kanso2012}. They cataloged these two integration as
\emph{Oracle and Reflection} and \emph{Oracle and Justification}, respectively.
Their integration with the \name{E} prover is the most related work with our
approach found at the moment while we write this document. Besides the use
of the same \ITP, crucial differences easily arise like decidability of
propositional logic using semantics that we do not take into account for the
reason we describe later on in section~\ref{sec:conclusions}.
